### 2011-09-13 paper/voisam/paper.tex

\documentclass{article}
\usepackage{algpseudocode}
\usepackage[ruled]{algorithm}
\usepackage{url}
\usepackage{framed}
\usepackage{amsfonts,amsmath,amsthm,amssymb}
\usepackage{graphicx}
\usepackage{url}
\usepackage{color}
\usepackage{geometry}

\geometry{margin=1.2in}

\newcommand {\mean} {\ensuremath {\mathop{\mathrm{mean}}}}
\newcommand {\median} {\ensuremath {\mathop{\mathrm{median}}}}
\newcommand {\N} {\ensuremath {\mathcal{N}}}
\newcommand {\IE} {\ensuremath {\mathbb{E}}}
\newcommand {\cov} {\ensuremath {\mathop{\mathrm{cov}}}}
\newcommand {\BEL} {\ensuremath {\mathop{\mathrm{BEL}}}}

\newtheorem{lemma}{Lemma}

\title{VOI-based Monte Carlo Sampling}
\author {David Tolpin, Solomon Eyal Shimony \\
Department of Computer Science, \\
Ben-Gurion University of the Negev, Beer Sheva, Israel \\
\{tolpin,shimony\}@cs.bgu.ac.il}

\begin{document}

\maketitle

\begin{abstract}
\end{abstract}

\section{Introduction and Definitions}

Taking a sequence of samples in order to minimize the
regret of a decision based on the samples is abstracted by the
{\em Multi-armed Bandit Problem.} In the Multi-armed Bandit problem
we have a set of $K$ arms. Each arm can be pulled multiple
times. When the $i$th arm is pulled, a random reward $X_i$ from an
unknown stationary distribution is returned.  The reward is bounded
between 0 and 1.

The simple regret of a sampling policy for the Multi-armed Bandit
Problem is the expected difference between the best expected reward
$\mu_*$ and the expected reward $\mu_j$ of the arm with the best sample mean
$\overline X_j=\max_i\overline X_i$:
\begin{equation}
\label{eq:simple-regret}
\IE[R]=\sum_{j=1}^K\Delta_j\Pr(\overline X_j=\max_i\overline X_i)
\end{equation}
where $\Delta_j=\mu_*-\mu_j$.
Strategies that minimize the simple regret are called pure exploration
strategies \cite{Bubeck.pure}. Principles of rational metareasoning
\cite{Russel.right} suggest that at each step the arm with the great
value of information (VOI) must be pulled, and the sampling must be
stopped and a decision must be made when no arm has positive VOI.

To estimate the VOI of pulling an arm, either a particular 
distribution of the rewards should be assumed (and updated based on
pull outcomes), or a distribution-independent bound on the VOI can be
used as the estimate. In this paper, we use {\em concentration inequalities}
to derive distribution-independent bounds on the VOI.

\section{Some Concentration Inequalities}

\begin{description}
\item[Markov' inequality \cite{}:] Let $X \ge 0$ be a random non-negative variable
  with finite mean $\IE[X]$. Then for all $a > 0$
\begin{equation}
\Pr(X\ge a)\le \frac {\IE[X]} a
\label{eq:conc-markov-upper}
\end{equation}
If additionally $X\le 1$, then by susbstituting $Y=1-X$
\begin{equation}
\Pr(X\le a)\le \frac {1-\IE[X]} {1-a}
\label{eq:conc-markov-lower}
\end{equation}
\item[Hoeffding's inequality \cite{}:] Let $X_1, \ldots, X_n$ be
independent random variables with values from $[0,1]$,
$X=\frac 1 n \sum_{i=1}^n X_i$. Then 
\begin{eqnarray}
\Pr(X-\IE[X] \ge a) &\le& \exp ( -2na^2)\nonumber\\
\Pr(\IE[X]-X \ge a) &\le& \exp ( -2na^2)
\label{eq:conc-hoeffding}
\end{eqnarray}
\item[Maurer \& Pontil's inequality \cite{MaurerPontil.benrstein}:] \footnote{A
  modification of Bennet's inequality \cite{} based on the sample,
  rather than the true, variance.}
\end{description}

\section{Upper Bounds on Value of Information}

\section{VOI-based Sampling Control}

\subsection{Selection Criterion}

\subsection{Termination Condition}

\bibliographystyle{plain}
\bibliography{refs}

\end{document}