### 2011-09-14 paper/voisam/paper.tex
added derivation for empirical bernstein inequality

\documentclass{article}
\usepackage{algpseudocode}
\usepackage[ruled]{algorithm}
\usepackage{url}
\usepackage{framed}
\usepackage{amsfonts,amsmath,amsthm,amssymb}
\usepackage{graphicx}
\usepackage{url}
\usepackage{color}
\usepackage{geometry}

\geometry{margin=1.2in}

\newcommand {\mean} {\ensuremath {\mathop{\mathrm{mean}}}}
\newcommand {\median} {\ensuremath {\mathop{\mathrm{median}}}}
\newcommand {\N} {\ensuremath {\mathcal{N}}}
\newcommand {\IE} {\ensuremath {\mathbb{E}}}
\newcommand {\cov} {\ensuremath {\mathop{\mathrm{cov}}}}
\newcommand {\BEL} {\ensuremath {\mathop{\mathrm{BEL}}}}

\newtheorem{lemma}{Lemma}

\title{VOI-based Monte Carlo Sampling}
\author {David Tolpin, Solomon Eyal Shimony \\
Department of Computer Science, \\
Ben-Gurion University of the Negev, Beer Sheva, Israel \\
\{tolpin,shimony\}@cs.bgu.ac.il}

\begin{document}

\maketitle

\begin{abstract}
\end{abstract}

\section{Introduction and Definitions}

Taking a sequence of samples in order to minimize the
regret of a decision based on the samples is abstracted by the
{\em Multi-armed Bandit Problem.} In the Multi-armed Bandit problem
we have a set of $K$ arms. Each arm can be pulled multiple
times. When the $i$th arm is pulled, a random reward $X_i$ from an
unknown stationary distribution is returned.  The reward is bounded
between 0 and 1.

The simple regret of a sampling policy for the Multi-armed Bandit
Problem is the expected difference between the best expected reward
$\mu_*$ and the expected reward $\mu_j$ of the arm with the best sample mean
$\overline X_j=\max_i\overline X_i$:
\begin{equation}
\label{eq:simple-regret}
\IE[R]=\sum_{j=1}^K\Delta_j\Pr(\overline X_j=\max_i\overline X_i)
\end{equation}
where $\Delta_j=\mu_*-\mu_j$.
Strategies that minimize the simple regret are called pure exploration
strategies \cite{Bubeck.pure}. Principles of rational metareasoning
\cite{Russel.right} suggest that at each step the arm with the great
value of information (VOI) must be pulled, and the sampling must be
stopped and a decision must be made when no arm has positive VOI.

To estimate the VOI of pulling an arm, either a certain 
distribution of the rewards should be assumed (and updated based on
observed rewards), or a distribution-independent bound on the VOI can be
used as the VOI estimate. In this paper, we use {\em concentration inequalities}
to derive distribution-independent bounds on the VOI.

\section{Some Concentration Inequalities}

Let $X_1, \ldots, X_n$ be i.i.d. random variables with values from $[0,1]$,
$X=\frac 1 n \sum_{i=1}^n X_i$. Then 
\begin{description}
\item[Hoeffding's inequality \rm{\cite{Hoeffding.ineq}}:] 
\begin{equation}
\Pr(X-\IE[X] \ge a) \le \exp ( -2na^2)
\label{eq:conc-hoeffding}
\end{equation}
\item[Empirical Bernstein's inequality
  \rm{\cite{MaurerPontil.benrstein}}:]\footnote{see
    Appendix~\ref{app:deriv-conc-empbernstein} for derivation}
\begin{eqnarray}
\Pr(X-\IE[X] \ge a) &\le& 2\exp \left( - \frac {na^2} {\frac {14} {3}
                          \frac {n} {n-1}a+2\overline\sigma_n^2}\right)\nonumber\\
                    &<& 2\exp \left( - \frac {na^2} {10a+2\overline\sigma_n^2}\right)
\label{eq:conc-empbernstein}
\end{eqnarray}
where sample variance $\overline\sigma_n^2$ is
\begin{equation}
\overline\sigma_n^2=\frac 1 {n(n-1)} \sum_{1\le i < j\le n}(X_i-X_j)^2
\label{eq:sample-variance}
\end{equation}
\end{description}
Bounds (\ref{eq:conc-hoeffding}, \ref{eq:conc-empbernstein}) are symmetrical
around the mean. Bound~(\ref{eq:conc-empbernstein}) is tighter than
(\ref{eq:conc-hoeffding}) for small $a$ and $\overline\sigma_n^2$.

\section{Upper Bounds on Value of Information}

\section{VOI-based Sampling Control}

\subsection{Selection Criterion}

\subsection{Termination Condition}

\appendix

\section{Empirical Bernstein Inequality}
\label{app:deriv-conc-empbernstein}

Theorem~4 in~\cite{MaurerPontil.benrstein} states that
\[\Pr\left(\IE[X]-\overline X_n \ge \sqrt { \frac {2\overline\sigma_n^2 \ln 2/\delta} n } + \frac {7 \ln 2/\delta} {3(n-1)}\right)\le \delta,\]
Therefore
\[\Pr\left(\IE[X]-\overline X_n \ge \sqrt { \left(\frac {7 \ln 2/\delta} {3(n-1)}\right)^2+\frac {2\overline\sigma_n^2 \ln 2/\delta} n } + \frac {7 \ln 2/\delta} {3(n-1)}\right)\le
\delta.\]
$t=\sqrt { \left(\frac {7 \ln 2/\delta} {3(n-1)}\right)^2+\frac {2\overline\sigma_n^2 \ln 2/\delta} n } + \frac {7 \ln 2/\delta} {3(n-1)}$ is a root of
square equation
\[t^2-t\frac {14 \ln 2/\delta} {3(n-1)} -\frac {2\overline\sigma_n^2 \ln 2/\delta} n=0\]
which, solved for $\delta\triangleq\Pr(\IE[X]-\overline X_n\ge t)$,
gives
\[\Pr(\IE[X]-\overline X_n\ge t)\le 2\exp \left( - \frac {na^2} {\frac {14} {3} \frac {n} {n-1}a+2\overline\sigma_n^2}\right)\]
Other derivations, giving slightly different results, are possible.

\bibliographystyle{plain}
\bibliography{refs}

\end{document}