### 2011-12-07 paper/voisam/onbetterbound.tex

\documentclass{article}
\usepackage{algpseudocode}
\usepackage[ruled]{algorithm}
\usepackage{url}
\usepackage{framed}
\usepackage{amsfonts,amsmath,amsthm,amssymb}
\usepackage{graphicx}
\usepackage{url}
\usepackage{color}
\usepackage{geometry}

\geometry{margin=1in}

\newcommand {\mean} {\ensuremath {\mathop{\mathrm{mean}}}}
\newcommand {\median} {\ensuremath {\mathop{\mathrm{median}}}}
\newcommand {\N} {\ensuremath {\mathcal{N}}}
\newcommand {\IE} {\ensuremath {\mathbb{E}}}
\newcommand {\cov} {\ensuremath {\mathop{\mathrm{cov}}}}
\newcommand {\BEL} {\ensuremath {\mathop{\mathrm{BEL}}}}

\newtheorem{lemma}{Lemma}

\title{On Better Bound on VOI of Monte Carlo Sampling}
\author {David Tolpin, Solomon Eyal Shimony \\
\{tolpin,shimony\}@cs.bgu.ac.il}

\begin{document}

\maketitle

Consider the case of VOI of measurement the current best item (sample mean
$\alpha$). The probability that the true mean of the item $\mu_\alpha$ is at
most $x$ is bounded by the Hoeffding's inequality as:
\[ \Pr(\mu_\alpha \le x) \le \exp\left(-2n(\alpha-x)^2\right) \]
An upper bound on the VOI is thus the product of the probability that
the true mean of the current best item is less than the sample mean
$\beta$ of the current second-best item times the maximum reward,
$\beta$:
\[ V_{\overline x=\alpha} \le \beta \exp\left(-2n(\alpha-\beta)^2\right) \]
The bound can be supposedly be improved by selecting a midpoint
$0 < \gamma < \beta$ and computing the bound as the sum of two parts:
\begin{itemize}
\item $\beta-\gamma$ multiplied by the probability that
  $\mu_\alpha \le \beta$;
\item $\beta$ multiplied by the probability that $\mu_\alpha\le
  \gamma$.
\end{itemize}.
\[ V_{\overline x=\alpha} \le (\beta-\gamma)\exp\left(-2n(\alpha-\beta)^2\right)+\beta\exp\left(-2n(\alpha-\gamma)^2\right) \]

\vspace{\baselineskip}

The minimum of $V^*$ is achieved when $\frac {dV^*} {d\gamma}=0$, that
is, when $\gamma$ is the root of the following equation:
\[ 4\beta n(\alpha-\gamma)=\exp\left(-2n\left(\frac {\alpha-\beta}
    {\alpha-\gamma}\right)^2\right) \]
If a root in the interval  $0\le\gamma\le\beta$ exists, then the
number of samples is bounded as
\[n\le\frac 1 {4\beta(\alpha-\beta)}\]
by observing that the right-hand side 
is at most $1$ (a negative power), and the left-hand side is at least
$4\beta n(\alpha-\beta)$.
So, the bound can supposedly be improved for smaller values of
$n$. The improvement is more significant when the current best and
second-best sample means are close.

The derivation for the other case
(sampling an item that can be better than the current best) is
obtained by substitution $1-\overline x, 1-\alpha, 1-\gamma$ instead of
$\alpha, \beta, \gamma$:

\[ V_{\overline x\ne\alpha} \le (\gamma-\alpha)\exp\left(-2n(\alpha-\overline
  x)^2\right)+(1-\alpha)\exp\left(-2n(\gamma-\overline x)^2\right) \]

The anticipated influence of the improved
VOI estimate would be that the selected item will be a less discovered
one and further from the current best or second-best.

A closed-form solution for $\gamma$ cannot be obtained, but given
$\alpha, \beta, n$, the value of $\gamma$ can be efficiently
computed. It should be determined empirically whether the improved
estimate has justified influence on the performance of the algorithm.

\end{document}