### 2012-02-26 paper/vago.ecai/paper.tex
short vago paper in progress

\documentclass{ecai2012}
\usepackage{times}
\usepackage{enumerate}
\usepackage{algpseudocode}
\usepackage[ruled]{algorithm}
\usepackage{url}
\usepackage{framed}
\usepackage{amsfonts,amsmath,amsthm,amssymb}
\usepackage{graphicx}
\usepackage{url}
\usepackage{color}

\ecaisubmission

\newcommand {\mean} {\ensuremath {\mathop{\mathrm{mean}}}}
\newcommand {\median} {\ensuremath {\mathop{\mathrm{median}}}}
\newcommand {\N} {\ensuremath {\mathcal{N}}}
\newcommand {\IE} {\ensuremath {\mathbb{E}}}
\newcommand {\cov} {\ensuremath {\mathop{\mathrm{cov}}}}
\newcommand {\BEL} {\ensuremath {\mathop{\mathrm{BEL}}}}

\newtheorem{dfn}{Definition}
\newtheorem{thm}{Theorem}
\newtheorem{lmm}{Lemma}
\newtheorem{crl}{Corollary}

\title{VOI-aware MCTS}
%\author {David Tolpin, Solomon Eyal Shimony \\
%Department of Computer Science, \\
%Ben-Gurion University of the Negev, Beer Sheva, Israel \\
%\{tolpin,shimony\}@cs.bgu.ac.il}

\begin{document}

\maketitle

\begin{abstract}
An MCTS sampling policy based on VOI estimates of rollouts is
suggested. Empirical evaluation of the policy and comparison to
UCB1 and UCT is performed on random Multi-armed Bandit instances
as well as on Computer Go.
\end{abstract}

\section{Introduction and Definitions}

TODO: start with MCTS, then go to Bandits.

Taking a sequence of samples in order to minimize the
regret of a decision based on the samples is abstracted by the
{\em Multi-armed Bandit Problem.} In the Multi-armed Bandit problem
we have a set of $K$ arms. Each arm can be pulled multiple
times. When the $i$th arm is pulled, a random reward $X_i$ from an
unknown stationary distribution is returned.  The reward is bounded
between 0 and 1.

The simple regret of a sampling policy for the Multi-armed Bandit
Problem is the expected difference between the best expected reward
$\max\limits_i\IE[X_i]$ and the expected reward $\IE[X_j]$ of the arm
with the best sample mean $\overline X_j=\max\limits_i\overline X_i$:
\begin{equation}
\label{eqn:simple-regret}
\IE[R]=\sum_{j=1}^K\Delta_j\Pr(\overline X_j=\max_i\overline X_i)
\end{equation}
where $\Delta_j=\max\limits_i\IE[X_i]-\IE[X_j]$.  Strategies that
minimize the simple regret are called pure exploration strategies
\cite{Bubeck.pure}. Principles of rational metareasoning
\cite{Russell.right} suggest that at each step the arm with the
greatest value of information (VOI) estimate must be pulled.

\section{Related Work}

Efficient algorithms for Multi-Armed Bandits based on
distribution-independent bounds, in particular UCB1, are introduced in
\cite{Auer.ucb}. The UCT algorithm, an extension of UCB1 to
MCTS is described in \cite{Kocsis.uct}, and a successful
application of UCT to playing Go is discussed in
\cite{Gelly.mogo}. Pure exploration in Multi-armed bandits is explored in
\cite{Bubeck.pure}. On the one hand, the paper proves certain upper
and lower bounds for UCB1 and uniform sampling, showing that an upper
bound on the simple regret is exponential in the number of samples for
uniform sampling, while only polynomial for UCB1. On the other hand,
empirical performance of UCB1 appears to be better than of uniform
sampling. The principles of bounded rationality appear in
\cite{Horvitz.reasoningabout}. \cite{Russell.right} provides a formal
description of rational metareasoning and case studies of applications
in several problem domains.

\section{Upper Bounds on Value of Information}

The intrinsic VOI $\Lambda_i$ of pulling an arm is the expected decrease
in the regret compared to selecting an arm without pulling any arm at
all. Two cases are possible:
\begin{itemize}
\item the arm $\alpha$ with the highest sample mean $\overline
  X_\alpha$ is pulled, and $\overline X_\alpha$ becomes lower than
  $\overline X_\beta$ of the second-best arm $\beta$;
\item another arm $i$ is pulled, and $\overline X_i$ becomes higher
than $\overline X_\alpha$.
\end{itemize}
The \textit{myopic} VOI estimate is of limited applicability to
Monte-Carlo sampling, since the effect of a single sample is small,
and the myopic VOI estimate will often be zero. However, $\Lambda_i$
can be estimated as the intrinsic value of information $\Lambda_i^b$
of pulling the $i$th arm for the rest of the budget of $N$ samples:
\begin{thm} Denote the current number of samples of the $i$th arm by
  $n_i$. $\Lambda_i^b$ is bounded from above as
\begin{align}
  \Lambda&_\alpha^b \le \frac {N \overline X_\beta} {N+n_i}\Pr(\overline X_i'\le\overline X_\beta)
    \le \frac {N \overline X_\beta} {n_i} \Pr(\overline X_i'\le\overline X_\beta)\nonumber\\
\Lambda&_{i|i\ne\alpha}^b \le \frac{ N(1-\overline  X_\alpha)} {N+n_i}\Pr(\overline X_i'\ge\overline X_\alpha)
     \le \frac {N(1-\overline X_\alpha)} {n_i}\Pr(\overline   X_i'\ge\overline X_\alpha)
\vspace{-24pt}
\label{eqn:thm-be}
\end{align}
where $\overline X_i'$ is the sample mean of the $i$th arm after $n_i+N$ samples.
\label{thm:be}
\end{thm}

The probabilities can be bounded from above using the
Hoeffding inequality \cite{Hoeffding.ineq}:
\begin{thm} The probabilities in equations (\ref{eqn:thm-be}) are bounded from above as
\begin{align}
\Pr&(\overline X_\alpha' \le \overline X_\beta)
  \le 2\exp\left(- \varphi(n_\alpha)(\overline X_\alpha - \overline X_\beta)^2 n_\alpha
   \right)\nonumber\\
\Pr&(\overline X_{i|i\ne\alpha}' \ge \overline X_\beta)
   \le 2\exp\left(- \varphi(n_i) (\overline X_\alpha -\overline  X_i)^2 n_i \right)
\label{eqn:probound-blnk-hoeffding}
\end{align}
where $\varphi(n)=2(\frac {1+n/N} {1+\sqrt {n/N}})^2 > 1.37$.
\label{thm:hoeffding-prob-bounds}
\end{thm}
\vspace{-16pt}
\begin{crl}
An upper bound on the VOI estimate $\Lambda_i^b$ is obtained
by substituting (\ref{eqn:probound-blnk-hoeffding}) into (\ref{eqn:thm-be}).
\begin{align}
  \Lambda&_\alpha^b \le \hat\Lambda_\alpha^b=\frac {2N\overline X_\beta} {n_\alpha}\exp\left(- 1.37(\overline X_\alpha - \overline X_\beta)^2 n_\alpha\right)\nonumber\\
  \Lambda&_{i|i\ne\alpha}^b\le \hat\Lambda_i^b=  \frac {2N(1-\overline  X_\alpha)} {n_i}\exp\left(- 1.37(\overline X_\alpha - \overline X_i)^2 n_i\right)
\label{eqn:bound-blnk-hoeffding}
\end{align}
\label{crl:bound-blnk-hoeffding}
\end{crl}
\vspace{-16pt}
Estimate $\hat\Lambda_i^b$ can be viewed as a
relaxation of the myopic assumption such that sequences consisting
of more than one action are taken into considertion.

\section{VOI-based Sample Allocation}

Following the principles of rational metareasoning, for pure
exploration in Multi-armed Bandits an arm with
the highest VOI should be pulled at each step. The upper bounds
established in Corollary~\ref{crl:bound-blnk-hoeffding} can be used
as VOI estimates. In MCTS, pure exploration takes place at the first
step of a rollout, where an action with the highest utility must
be chosen. MCTS differs from pure epxloration in Multi-armed Bandits
in that the distributions of the rewards are not stationary. However,
VOI estimates computed as for stationary distributions work well in
practice. As illustrated by the empirical evaluation
(Section~\ref{sec:empirical-evaluation}), estimates based on upper
bounds on the VOI result in a rational sampling policy exceeding the
performance of some state-of-the-art heuristic algorithms.

\section{Empirical Evaluation}
\label{sec:empirical-evaluation}

\subsection{Selecting The Best Arm}
\label{sec:emp-arm}

\begin{figure}[h!]
\centering
\includegraphics[scale=0.5]{flat.pdf}
\vspace{-8pt}
\caption{Random instances: regret vs. number of samples}
\label{fig:random-instances}
\vspace{-16pt}
\end{figure}

The sampling policies are first compared on random Multi-armed bandit 
problem instances. Figure~\ref{fig:random-instances} shows results for
randomly-generated Multi-armed bandits with 32 Bernoulli arms, with
the mean rewards of the arms distributed uniformly in the range~$[0,
  1]$, for a range of sample budgets~$32..1024$, with multiplicative
step of~$2$. The experiment for each number of samples was repeated
10000 times. UCB1 is always considerably worse than the
VOI-aware sampling policy.

\subsection{Playing Go Against UCT}
\label{sec:emp-go}

The policies were also compared on Computer Go, a  search domain
in which UCT-based MCTS has been particularly successful
\cite{Gelly.mogo}. A modified version of Pachi \cite{Braudis.pachi}, a state of the art
Go program, was used for the experiments. The UCT engine was extended
with a VOI-aware sampling policy, and a time allocation mode ensuring
that both the original UCT policy and the VOI-aware policy use the
same average number of samples per node was added. (While the UCT
engine is not the most powerful engine of Pachi, it is still a strong
player; on the other hand, additional features of more advanced
engines would obstruct the MCTS phenomena which are the subject of
the experiment.)
\begin{figure}[h!]
\centering
\includegraphics[scale=0.5]{vct-wins.pdf}
\vspace{-8pt}
\caption{Go: winning rate --- UCT against VCT}
\label{fig:vct-against-uct}
\vspace{-16pt}
\end{figure}
The engines were compared on the 9x9 board, for 5000, 7000, 1000, and
15000 samples per ply, each experiment was repeated
1000 times. Figure~\ref{fig:vct-against-uct}
shows the winning rate of VOI+UCT against UCT vs. the number of
samples. For most numbers of samples per node, VOI+UCT outperforms UCT.

\section{Summary and Future Work}

This work suggested a Monte-Carlo sampling policy in which sample
selection is based on upper bounds on the value of
information. Empirical evaluation showed that this policy outperforms
heuristic algorithms for pure exploration in Multi-armed
bandits, as well as for MCTS.

MCTS still remains a largely unexplored field of
application of VOI-aware algorithms. More elaborated VOI estimates,
taking into consideration re-use of samples in future search states
should be considered. The policy introduced in the paper differs from
the UCT algorithm only at the first step, where the VOI-aware
decisions are made. Consistent application of principles of rational
metareasoning at all steps of a rollout may further improve the
sampling.

\bibliographystyle{ecai2012}
\bibliography{refs}

\end{document}