### 2012-03-20 paper/vago.uai/paper.tex
vago.uai in progress

\documentclass[]{article}
\usepackage{proceed2e}
\usepackage{enumerate}
\usepackage{algpseudocode}
\usepackage[ruled]{algorithm}
\usepackage{url}
\usepackage{framed}
\usepackage{amsfonts,amsmath,amsthm,amssymb}
\usepackage{graphicx}
\usepackage{url}
\usepackage{color}

\newcommand {\mean} {\ensuremath {\mathop{\mathrm{mean}}}}
\newcommand {\median} {\ensuremath {\mathop{\mathrm{median}}}}
\newcommand {\N} {\ensuremath {\mathcal{N}}}
\newcommand {\IE} {\ensuremath {\mathbb{E}}}
\newcommand {\cov} {\ensuremath {\mathop{\mathrm{cov}}}}
\newcommand {\BEL} {\ensuremath {\mathop{\mathrm{BEL}}}}

\newtheorem{dfn}{Definition}
\newtheorem{thm}{Theorem}
\newtheorem{lmm}{Lemma}
\newtheorem{crl}{Corollary}

\begin{document}

\section*{Upper Bounds on Value of Information}

The intrinsic VOI $\Lambda_i$ of pulling an arm is the expected decrease
in the regret compared to selecting the best arm without pulling any arm at
all. Two cases are possible:
\begin{itemize}
\item the arm $\alpha$ with the highest sample mean $\overline
  X_\alpha$ is pulled, and $\overline X_\alpha$ becomes lower than
  $\overline X_\beta$ of the second-best arm $\beta$;
\item another arm $i$ is pulled, and $\overline X_i$ becomes higher
than $\overline X_\alpha$.
\end{itemize}
The \textit{myopic} VOI estimate is of limited applicability to
Monte-Carlo sampling, since the effect of a single sample is small,
and the myopic VOI estimate will often be zero. However, for the
common case of a fixed budget of samples per node, $\Lambda_i$ can be
estimated as the intrinsic VOI $\Lambda_i^b$ of pulling the $i$th arm
for the rest of the budget.  Let us denote the current number of
samples of the $i$th arm by $n_i$, and the remaining number of samples
by $N$:
\begin{thm} $\Lambda_i^b$ is bounded from above as
\begin{eqnarray}
\label{eqn:thm-be}
  &\Lambda_\alpha^b \le \frac {N \overline X_\beta} {N+n_i}\Pr(\overline X_i'\le\overline X_\beta)
    \le \frac {N \overline X_\beta} {n_i} \Pr(\overline X_i'\le\overline X_\beta)\\
&\Lambda_{i|i\ne\alpha}^b \le \frac{ N(1-\overline  X_\alpha)} {N+n_i}\Pr(\overline X_i'\ge\overline X_\alpha)
     \le \frac {N(1-\overline X_\alpha)} {n_i}\Pr(\overline   X_i'\ge\overline X_\alpha)\nonumber
\end{eqnarray}
where $\overline X_i'$ is the sample mean of the $i$th arm after $n_i+N$ samples.
\label{thm:be}
\end{thm}

The probabilities can be bounded from above using the
Hoeffding inequality \cite{Hoeffding.ineq}:
\begin{thm} The probabilities in equations (\ref{eqn:thm-be}) are bounded from above as
\begin{align}
  \label{eqn:probound-blnk-hoeffding}
  \Pr&(\overline X_\alpha' \le \overline X_\beta)
  \le 2\exp\left(- \varphi(n_\alpha)(\overline X_\alpha - \overline X_\beta)^2 n_\alpha
  \right)\nonumber\\
  \Pr&(\overline X_{i|i\ne\alpha}' \ge \overline X_\beta)
  \le 2\exp\left(- \varphi(n_i) (\overline X_\alpha -\overline  X_i)^2 n_i \right)
\end{align}
where $\varphi(n)=2(\frac {1+n/N} {1+\sqrt {n/N}})^2 > 1.37$.
\label{thm:hoeffding-prob-bounds}
\end{thm}
\begin{crl}
An upper bound on the VOI estimate $\Lambda_i^b$ is obtained
by substituting (\ref{eqn:probound-blnk-hoeffding}) into (\ref{eqn:thm-be}).
\begin{align}
  \label{eqn:bound-blnk-hoeffding}
  \Lambda&_\alpha^b \le \hat\Lambda_\alpha^b=\frac {2N\overline X_\beta} {n_\alpha}\exp\left(- 1.37(\overline X_\alpha - \overline X_\beta)^2 n_\alpha\right)\nonumber\\
  \Lambda&_{i|i\ne\alpha}^b\le \hat\Lambda_i^b=  \frac {2N(1-\overline  X_\alpha)} {n_i}\exp\left(- 1.37(\overline X_\alpha - \overline X_i)^2 n_i\right)
\end{align}
\label{crl:bound-blnk-hoeffding}
\end{crl}

\section*{VOI-based Sample Allocation}

Following the principles of rational metareasoning, for pure
exploration in Multi-armed Bandits an arm with
the highest VOI should be pulled at each step. The upper bounds
established in Corollary~\ref{crl:bound-blnk-hoeffding} can be used
as VOI estimates. In MCTS, pure exploration takes place at the first
step of a rollout, where an action with the highest utility must
be chosen. MCTS differs from pure exploration in Multi-armed Bandits
in that the distributions of the rewards are not stationary. However,
VOI estimates computed as for stationary distributions work well in
practice. As illustrated by the empirical evaluation
(Section~\ref{sec:empirical-evaluation}), estimates based on upper
bounds on the VOI result in a rational sampling policy exceeding the
performance of some state-of-the-art heuristic algorithms.

\section*{Empirical Evaluation}
\label{sec:empirical-evaluation}

\subsection*{Selecting The Best Arm}
\label{sec:emp-arm}

\begin{figure}[h!]
\centering
\includegraphics[scale=0.6]{flat.pdf}
\vspace{-8pt}
\caption{Random instances: regret vs. number of samples}
\label{fig:random-instances}
\vspace{-16pt}
\end{figure}

The sampling policies are first compared on random Multi-armed bandit 
problem instances. Figure~\ref{fig:random-instances} shows results for
randomly-generated Multi-armed bandits with 32 Bernoulli arms, with
the mean rewards of the arms distributed uniformly in the range~$[0,
  1]$, for a range of sample budgets~$32..1024$, with multiplicative
step of~$2$. The experiment for each number of samples was repeated
10000 times. UCB1 is always considerably worse than the
VOI-aware sampling policy.

\subsection*{Playing Go Against UCT}
\label{sec:emp-go}

The policies were also compared on Computer Go, a  search domain
in which UCT-based MCTS has been particularly successful
\cite{Gelly.mogo}. A modified version of Pachi \cite{Braudis.pachi}, a state of the art
Go program, was used for the experiments. The UCT engine was extended
with a VOI-aware sampling policy, and a time allocation mode ensuring
that both the original UCT policy and the VOI-aware policy use the
same average number of samples per node was added. (While the UCT
engine is not the most powerful engine of Pachi, it is still a strong
player; on the other hand, additional features of more advanced
engines would obstruct the MCTS phenomena which are the subject of
the experiment.)
\begin{figure}[h!]
\centering
\includegraphics[scale=0.6]{vct-wins.pdf}
\vspace{-8pt}
\caption{Go: winning rate --- VOI against UCT}
\label{fig:vct-against-uct}
\vspace{-16pt}
\end{figure}
The engines were compared on the 9x9 board, for 5000, 7000, 10000, and
15000 samples per ply, each experiment was repeated
1000 times. Figure~\ref{fig:vct-against-uct}
shows the winning rate of VOI against UCT vs. the number of
samples. For most numbers of samples per node, VOI outperforms UCT.

\bibliographystyle{plain}
\bibliography{refs}

\end{document}