### 2012-03-20 paper/vago.uai/uai12-flat.tex

\documentclass{article}
\begin{document}

\section{Introduction}

%% Build on the intro from the Hay and Russell tech report

[[Sequential decision problem approximation by simulating possible futures]]

[[Metareasoning to select simulations]]

[[Monte-Carlo tree search]]

[[UCT, which is based on the asymptotically regret-optimal bandit control rule UCB]]

[[Bandit problems vs. selection problems]]

[[Resulting problems for UCT: biased against selecting negative options, no natural stopping criteria]]

[[explain nature of our results: basic theoretical foundations,
    initial foray into new hueristics for selection and some empirical
    results]]

\section{On optimal policies for selection}

[[Formal definition of selection problems and the metalevel MDP with cost per sample (time value); also mention budgeted learning.]]

[[Theorem: if Optimal stops in x, myopic stops in x (converse is more useful!)]]

[[Theorem: if Myopic stops in all states reachable from x, optimal stops in x]]

[[Theorem: expected number of samples is bounded; actual number bounded wp 1]]

[[Theorems: actual number of samples bounded for flat Normal, Bernoulli]]

[[Counterexample to actual boundedness in general (SPRT case)]]

\section{Context effects and non-indexability}

[[No index theorem; via context inversion counter-example]]

[[Theorem: context effect occurs only for a single convex interval of
    context value (how general can we make this?]]

\section{Regret bounds and approximate policies}

[[Regret models: simple regret, regret with cost per sampling; regret goes to zero as c does]]

[[Expected simple regret bounds for normal case?]]

[[Blinkered sampling]]

[[ESPb: Frazier's continuous time approximation]]

%% [[Selection criteria]]
%% [[Termination criteria]]

[[Concentration upper bounds on VOI]]

[[Control rules: VCT, ECT, BCT]]

\section{Application to Monte Carlo tree search}

[[summarize MCTS]]

[[Root rules vs. subtree rules; here just focus on former.]]

[[Efficiency gains from reusing samples between time steps; Shimony's model of this]]

[[Simple regret in bandit problem [VEB]CT vs. UCB]]

[[Improvement in performance in Go (or other trees) [VEB]CT vs. UCT]]

\end{document}