### 2012-03-21 paper/vago.uai/paper.tex
draft sent

\documentclass[]{article}
\usepackage{proceed2e}
\usepackage{enumerate}
\usepackage{algpseudocode}
\usepackage[ruled]{algorithm}
\usepackage{url}
\usepackage{framed}
\usepackage{amsfonts,amsmath,amsthm,amssymb}
\usepackage{graphicx}
\usepackage{url}
\usepackage{color}

\newcommand {\mean} {\ensuremath {\mathop{\mathrm{mean}}}}
\newcommand {\median} {\ensuremath {\mathop{\mathrm{median}}}}
\newcommand {\N} {\ensuremath {\mathcal{N}}}
\newcommand {\IE} {\ensuremath {\mathbb{E}}}
\newcommand {\cov} {\ensuremath {\mathop{\mathrm{cov}}}}
\newcommand {\BEL} {\ensuremath {\mathop{\mathrm{BEL}}}}

\newtheorem{dfn}{Definition}
\newtheorem{thm}{Theorem}
\newtheorem{lmm}{Lemma}
\newtheorem{crl}{Corollary}

\begin{document}

\section{Upper Bounds on Value of Information}

The intrinsic (myopic greedy) VOI $\Lambda_i$ of testing an arm $i$ is
the expected reward of the best arm given the information obtained by the test,
minus the expected reward of the best arm according to the current information state.
We will assume here a distribution model where the arm with the highest sample mean
in an information state is the one with the highest expected reward.
Two cases are possible:
\begin{itemize}
\item the arm $\alpha$ with the highest sample mean $\overline
  X_\alpha$ is tested, and $\overline X_\alpha$ becomes lower than
  $\overline X_\beta$ of the second-best arm $\beta$;
\item another arm $i$ is tested, and $\overline X_i$ becomes higher
than $\overline X_\alpha$.
\end{itemize}

In order to estimate the VOI of testing an arm, a sampling model is
needed.  The generic sampling model used is: 
\begin{itemize}
\item The actual value of an arm has a certain (possibly unknown) distribution.
\item Testing an arm is an i.i.d. noisy measurement of the actual
  value of the arm.
\end{itemize}
When the distribution model is known, one can use the methods of Section \ref{sec:normal}.
Otherwise, one can still use {\em concentration
inequalities} to derive distribution-independent bounds on the
VOI. Our bounds are applicable to any bounded distribution. Similar
bounds can be derived for certain unbounded distributions, such as the
normally distributed prior of the value with normally distributed
sampling.

Unfortunately, the pure \textit{myopic} VOI estimate is of little use in
Monte-Carlo sampling, since the effect of a single sample is small,
and the myopic VOI estimate will often be zero. Instead, we will derive
an estimate for the case where we decide to test an arm $N$ times,
where $N$ can be the remaining ``budget'' of available samples or
any other integer quantity.
Denote by  $\Lambda_i^b$ the intrinsic VOI of testing the $i$th arm
$N$ times, and the number of
samples already taken from the $i$th arm by $n_i$.
\begin{thm} $\Lambda_i^b$ is bounded from above as
\begin{align}
\label{eqn:thm-be}
  \Lambda_\alpha^b&\le \frac {N \overline X_\beta^{n_\beta}} {n_\alpha} \Pr(\overline X_\alpha^{n_\alpha+N}\le\overline X_\beta^{n_\beta})\nonumber\\
\Lambda_{i|i\ne\alpha}^b&\le \frac {N(1-\overline X_\alpha^{n_\alpha})} {n_i}\Pr(\overline   X_i^{{n_i}+N}\ge\overline X_\alpha^{n_\alpha})
\end{align}
\label{thm:be}
\end{thm}
\vspace{-2em}
\begin{proof} For the case $i\ne \alpha$, the probability that the
  $i$th arm is finally chosen instead of $\alpha$ is
  $\Pr(\overline X_i^{n_i+N} \ge \overline X_\alpha^{n_\alpha})$. $X_i \le 1$,
  therefore $\overline X_i^{n_i+N}\le \overline
  X_\alpha^{n_\alpha}+\frac {N(1-\overline X_\alpha^{n_\alpha})} {N+n_i}$. Hence, the intrinsic value of perfect
  information is at most: 
\begin{equation}
\label{eq:simplistic}
\frac{ N(1-\overline  X_\alpha^{n_\alpha})}
  {N+n_i}\Pr(\overline X_i^{{n_i}+N}\hspace{-0.5em}\ge\overline X_\alpha^{n_\alpha})\le \frac {N \overline X_\beta^{n_\beta}} {n_\alpha} \Pr(\overline X_\alpha^{n_\alpha+N}\hspace{-0.5em}\le\overline X_\beta^{n_\beta})
\end{equation}
  Proof for the case $i=\alpha$ is similar.
\end{proof}

The probabilities can be bounded from above using the
Hoeffding inequality \cite{Hoeffding.ineq}:
\begin{thm} The probabilities in equations (\ref{eqn:thm-be}) are bounded from above as
\begin{align}
  \label{eqn:probound-blnk-hoeffding}
  \Pr&(\overline X_\alpha^{{n_\alpha}+N} \le \overline X_\beta^{n_\beta})
  \le 2\exp\left(- \varphi (\overline X_\alpha^{n_\alpha} - \overline X_\beta^{n_\beta})^2 n_\alpha
  \right)\nonumber\\
  \Pr&(\overline X_{i|i\ne\alpha}^{n_\alpha+N} \ge \overline X_\beta^{n_\beta})
  \le 2\exp\left(- \varphi (\overline X_\alpha^{n_\alpha} -\overline  X_i^{n_i})^2 n_i \right)
\end{align}
where $\varphi=\min \left(2(\frac {1+n/N} {1+\sqrt {n/N}})^2\right)=8(\sqrt 2 - 1)^2 > 1.37$.
\label{thm:hoeffding-prob-bounds}
\end{thm}
\begin{proof}
Equations (\ref{eqn:probound-blnk-hoeffding}) follow from the
observation that if $i\ne\alpha$, $\overline X_i^{n_i+N}>\overline X_\alpha^{n_i}$
if and only if the mean $\overline X_i^N$ of $N$ samples from $n_i+1$
to $n_i+N$ is at least $\overline X_\alpha^{n_i}+(\overline X_\alpha^{n_i}-\overline
X_i^{n_i})\frac {n_i} N$.

For any $\delta$, the probability that $\overline X_i^{n_i+N}$ is greater
than $\overline X_\alpha^{n_i}$ is less than the probability that
$\IE[X_i]\ge\overline X_i^n+\delta$
\emph{or} $\overline X_i^N\ge \IE[X_i]+\overline X_\alpha^{n_\alpha}
- \overline X_i^{n_i} - \delta +(\overline X_\alpha^{n_\alpha} - \overline X_i^{n_i})\frac {n_i} N$,
thus, by the union bound, less than the sum of the probabilities:
\begin{align}
\Pr&(\overline X_i^{n_i+N}\ge \overline X_\alpha^{n_i})\nonumber\\
   &\le\Pr(\IE[X_i]-\overline X_i^{n_i} \ge \delta)\\
   &\hspace{0.25em}+\Pr\left(\overline X_i^N\hspace{-0.5em} - \IE[X_i] \ge \overline X_\alpha^{n_\alpha}\hspace{-0.5em}
           - \overline X_i^{n_i}\hspace{-0.5em} - \delta +(\overline X_\alpha^{n_\alpha}\hspace{-0.5em} - \overline X_i^{n_i})\frac {n_i} N\right)\nonumber
\end{align}
Bounding the probabilities on the right-hand side using the Hoeffding
inequality, obtain:
\begin{align}
\Pr&(\overline X_i^{n_i+N}\ge \overline X_\alpha^{n_\alpha})\le \nonumber\\
    &\exp(-2\delta^2n_i)+\nonumber\\
    &\quad\exp\left(-2\left((\overline X_\alpha^{n_\alpha}\hspace{-0.5em}
         - \overline X_i^{n_i})\left(1+\frac {n_i} N\right)
         - \delta\right)^2N\right)
\label{eqn:app-hoeffding-le-maxexp}
\end{align}
Find $\delta$ for which the two terms on the right-hand side of
(\ref{eqn:app-hoeffding-le-maxexp}) are equal:
\begin{equation}
\exp(-\delta^2n) = \exp\left(-2\left((\overline X_\alpha - \overline X_i)(1+\frac {n_i} N) - \delta\right)^2N\right)\label{eqn:app-hoeffding-eq-exp}
\end{equation}
Solve (\ref{eqn:app-hoeffding-eq-exp}) for $\delta$: $\delta=\frac {1+\frac {n_i} N} {1+\sqrt {\frac {n_i} N}} (\overline X_\alpha^{n_\alpha}\hspace{-0.5em}
- \overline X_i^{n_i}) \ge 2(\sqrt 2 - 1)(\overline X_\alpha^{n_\alpha}\hspace{-0.5em}-\overline X_i^{n_i})$. Substitute $\delta$ into 
(\ref{eqn:app-hoeffding-le-maxexp}) and obtain
\begin{align}
\Pr&(\overline X_i^{n_i}\ge \overline X_\alpha^{n_\alpha}) \nonumber\\
& \le 2\exp\left(-2\left( \frac {1+\frac {n_i} N} {1+\sqrt {\frac {n_i} N}}
                          (\overline X_\alpha^{n_\alpha} - \overline X_i^{n_i})\right)^2 n_i\right)\nonumber \\
& \le 2\exp(-8(\sqrt 2 - 1)^2(\overline X_\alpha^{n_\alpha} - \overline X_i^{n_i})^2n_i)\nonumber\\
& = 2\exp(-\varphi(\overline X_\alpha^{n_\alpha} - \overline X_i^{n_i})^2n_i)
\end{align}
Derivation for the case $i=\alpha$ is similar.
\end{proof}

\begin{crl}
An upper bound on the VOI estimate $\Lambda_i^b$ is obtained
by substituting (\ref{eqn:probound-blnk-hoeffding}) into (\ref{eqn:thm-be}).
\begin{align}
  \label{eqn:bound-blnk-hoeffding}
  \Lambda&_\alpha^b \le \hat\Lambda_\alpha^b=\frac {2N\overline X_\beta^{n_\beta}} {n_\alpha}\exp\left(- \varphi(\overline X_\alpha^{n_\alpha}\hspace{-0.5em} - \overline X_\beta^{n_\beta})^2 n_\alpha\right)\nonumber\\
  \Lambda&_{i|i\ne\alpha}^b\le \hat\Lambda_i^b=  \frac {2N(1-\overline  X_\alpha^{n_\alpha})} {n_i}\exp\left(- \varphi(\overline X_\alpha^{n_\alpha}\hspace{-0.5em} - \overline X_i^{n_i})^2 n_i\right)
\end{align}
\label{crl:bound-blnk-hoeffding}
\end{crl}
\vspace{-2em}

More refined bounds can be obtained through tighter estimates on the
probabilities in Equation \ref{eq:simplistic}, for example, based on the empirical Bernstein
inequality~\cite{MaurerPontil.benrstein}, or through a more careful
application of the Hoeffding inequality, resulting in:
\begin{align}
\Lambda_i^b&\le\frac {N\sqrt \pi} {n_i \sqrt {n_i}}
  \left[\mathrm{erf}\left((1\hspace{-0.25em}-\hspace{-0.25em}\overline X_i^{n_i})\sqrt {n_i}\right)
      \hspace{-0.25em}-\hspace{-0.25em}\mathrm{erf}\left((\overline X_\alpha^{n_\alpha}\hspace{-0.5em} - \overline X_i^{n_i})\sqrt{n_i}\right)\right]\nonumber\\
\Lambda_\alpha^b&\le\frac {N\sqrt \pi} {n_\alpha \sqrt {n_\alpha}}
  \left[\mathrm{erf}\left(\overline X_\alpha^{n_\alpha}\sqrt {n_\alpha}\right)
      -\mathrm{erf}\left((\overline X_\alpha^{n_\alpha}\hspace{-0.5em} - \overline X_\beta^{n_\beta})\sqrt{n_\alpha}\right)\right]
\label{eqn:erf-blinkered}
\end{align}

\section{VOI-based Sampling Control}

\subsection{Selection Criterion}

A reasonable greedy policy is to test the arm that has the best
current VOI estimate. When the distributions are unknown, it makes sense
to use the upper bounds established in Theorem~\ref{thm:be} instead. Although MCTS
differs from pure exploration in Multi-armed Bandits in that the
distributions of the rewards are not stationary, VOI
estimates computed as for stationary distributions work well in
practice. As illustrated
by the empirical evaluation (Section~\ref{sec:empirical-evaluation}),
estimates based on upper bounds on the VOI result in good sampling
policies, which exhibit performance comparable to the performance of
some state-of-the-art heuristic algorithms.

\subsection{Stopping Criterion}
\label{sec:control-stopping-criterion}

The simplest stopping criterion for a sampling policy is the
budget---a fixed number of samples performed before a decision is
made. When a sample has a known cost commensurable with the value of
information of a measurement, an upper bound on the intrinsic VOI can also
be used to stop the sampling if the intrinsic VOI of any action
is less than the cost of sampling $C$:
\begin{equation}
\mbox{stop if } \max_i \Lambda_i \le C
\end{equation}
VOI estimates~(\ref{eqn:thm-be},
\ref{eqn:bound-blnk-hoeffding}) include the remaining sample budget $N$ as a
factor, but given the cost of a single sample $c$, the cost of the
remaining samples accounted for in estimating the intrinsic VOI is
$C=cN$. $N$ can be dropped on both sides of the inequality,
and a reasonable stopping criterion would be:
\begin{align}
\frac 1 N \Lambda_\alpha^b \le&\frac {\overline X_\beta^{n_\beta}}
  {n_\alpha}\Pr(\overline X_\alpha^{n_\alpha+N}\le\overline
  X_\beta^{n_\alpha})\le c\nonumber\\
\frac 1 N \max_i\Lambda_i^b\le &\max_i\frac {(1-\overline X_\alpha^{n_\alpha})} {n_i}\Pr(\overline
  X_i^{n_i+N}\ge\overline X_\alpha^{n_\alpha})\le c\nonumber\\
    &\forall i: i\ne\alpha
\label{eqn:stopping-blnk}
\end{align}
The empirical evaluation (Section~\ref{sec:empirical-evaluation})
confirms the viability of this stopping criterion and illustrates the
influence of the sample cost $c$ on the performance of
the sampling policy.

\subsection{Sample Redistribution in Trees}
\label{sec:control-redistribution}

Monte-Carlo tree search \cite{Chaslot.montecarlo} selects at each
search step an action that appears to be the best based on outcomes
of \textit{search rollouts}. Two different criteria are employed in
action selection during a rollout:
\begin{itemize}
\item at the first step, the value of partial
information should be used to select nodes to sample, as the goal is
selection of a move at the current root node;
\item as the goal of sampling in deeper tree nodes is usually to improve the
node value estimate, rather than selection, a different sampling scheme
should be used there. At these levels, the standard practice
of using UCT-like rollouts makes sense, as this method is known to be nearly optimal
in minimizing the \textit{cumulative} regret.
\end{itemize}

The above hybrid approach assumes
that the information obtained from rollouts in the
current state is discarded after an action is selected. In practice,
many successful Monte-Carlo tree search algorithms re-use rollouts
generated at earlier search states, if the sample traverses the
current search state during the rollout; thus, the value of information of a rollout is
determined not just by the influence on the choice of the action at
the current state, but also by its potential influence on the choice at future
search states.

One way to account for this re-use would be to incorporate the
`future' value of information into a VOI estimate. However, this 
approach requires a non-trivial extension of the theory of meta-reasoning for search.
Alternately, one can behave myopically w.r.t. search tree depth:
\begin{enumerate}
\item estimate VOI as though the information is discarded after each step;
\item stop early if the VOI is below a certain threshold
   (see Section~\ref{sec:control-stopping-criterion});
\item save the unused sample budget for search in future states, such that
   if the nominal budget is $N$, and the unused budget in the last state
   is $N_u$, the search budget in the next state will be $N+N_u$.
\end{enumerate}
In this approach, the cost of a sample in the current state is the
VOI of increasing the budget of a future state by one sample.  It is
unclear whether this cost can be accurately estimated, but supposedly
a fixed value for a given problem type and algorithm implementation
would work. Indeed, the empirical evaluation (Section~\ref{sec:emp-go})
confirms that stopping and sample redistribution based on a learned
fixed cost  substantially improve the performance of the VOI-based
sampling policy in game tree search.

\section{Empirical Evaluation}
\label{sec:empirical-evaluation}

\subsection{Selecting The Best Arm}
\label{sec:emp-arm}

\begin{figure}[h!]
\centering
\includegraphics[scale=0.6]{flat.pdf}
\caption{Random instances: regret vs. number of samples}
\label{fig:random-instances}
\end{figure}

The sampling policies are first compared on random Multi-armed bandit 
problem instances. Figure~\ref{fig:random-instances} shows results for
randomly-generated Multi-armed bandits with 25 Bernoulli arms, with
the mean rewards of the arms distributed uniformly in the range~$[0,
  1]$, for a range of sample budgets~$200..2000$, with multiplicative
step of~$2$. The experiment for each number of samples was repeated
10000 times. UCB1, as well as policies  based on bounds
(\ref{eqn:bound-blnk-hoeffding}) (VOI) and
(\ref{eqn:erf-blinkered}) (VOI+) are compared. UCB1 is always considerably worse than the
VOI-aware sampling policies.

\subsection{Playing Go Against UCT}
\label{sec:emp-go}

The policies were also compared on Computer Go, a  search domain
in which UCT-based MCTS has been particularly successful
\cite{Gelly.mogo}. A modified version of Pachi \cite{Braudis.pachi}, a state of the art
Go program, was used for the experiments:
\begin{itemize}
\item The UCT engine of Pachi was extended with VOI-aware sampling
  policies at the first step. 
\item The stopping criterion for the VOI-aware policy was
  modified and based solely on the sample cost, specified as
  a constant parameter. The heuristic stopping criterion for the
  original UCT policy was left unchanged.
\item The time-allocation mode based on the fixed number of samples
  was modified for \textit{both the original UCT policy and the VOI-aware
  policies} such that 
  \begin{itemize}
    \item the same number of samples is available to
      the agent at each step, independently of the number of pre-simulated
      games;  
    \item if samples were unused at the current step,
      they become available at the next step. 
  \end{itemize}
\end{itemize}
While the UCT engine is not the most powerful engine of Pachi, it is still a strong
player. On the other hand, additional features of more advanced
engines would obstruct the MCTS phenomena which are the subject of
the experiment.
\begin{figure}[h!]
\centering
\includegraphics[scale=0.6]{uctvoi.pdf}
\caption{Go: winning rate vs. $c$}
\label{fig:uctvoi}
\end{figure}
The engines were compared on the 9x9 board, for 5000, 7000, 1000, and
15000 samples (game simulations) per ply, each experiment was repeated
1000 times. Figure~\ref{fig:uctvoi}
show the winning rate of the VOI-aware policy against UCT
vs. the stopping threshold $c$ (if the maximum VOI of a sample is below
the threshold, the simulation is stopped, and a move is chosen). Each
curve in the figure corresponds to a certain number of samples per
ply. For the stopping threshold of $10^{-6}$, the VOI-aware policy
is almost always better than UCT, and  reaches the winning rate of
64\% for 10000 samples per ply.

\begin{figure}[h!]
\centering
\includegraphics[scale=0.6]{voi-wins.pdf}
\caption{Go: winning rate for $c=10^{-6}$}
\label{fig:voi-wins}
\end{figure}

Figure~\ref{fig:voi-wins}
shows the winning rate of VOI against UCT vs. the number of
samples for $c=10^{-6}$. In agreement with the intuition
(Section~\ref{sec:control-redistribution}), VOI-based stopping and
sample redistribution is most influential for intermediate numbers of
samples per ply. When the maximum number of samples is too low, early
stopping would result in poorly selected moves. On the other hand,
when the maximum number of samples is sufficiently high, the VOI of
increasing the maximum number of samples in a future state is low.

\bibliographystyle{plain}
\bibliography{refs}

\end{document}