### 2012-03-28 paper/vago.uai/introduction.tex

The broad family of sequential decision problems includes
combinatorial search problems, game playing, robotic path planning,
model-predictive control problems, Markov decision processes (fully or
partially observable), and a huge range of applications. In almost all
realistic instances, exact solution is intractable and approximate
methods are sought.  Perhaps the most popular approach is to simulate
a limited number of possible future action sequences. For example,
a typical game-playing algorithm explores a tree or graph of action sequences
with some limit on depth, using pruning methods to avoid irrelevant
subtrees; based on what it finds in the explored portion of the state space, the algorithm
then selects a move.

Clearly, it is desirable to select the best possible move from the
least possible amount of exploration. For a given amount of
exploration, decision quality can be improved by directing exploration
towards those actions sequences whose outcomes are helpful in selecting
a good move. Thus, the {\em metalevel} decision problem is to choose
what future action sequences to explore (or, more generally, what
deliberative computations to do), while the {\em object-level}
decision problem is to choose an action to execute in the real world.

That the metalevel decision problem can itself be formulated and
solved decision-theoretically was noted by \citet{Matheson:1968},
borrowing directly from the related concept of {\em information value
  theory}~\citep{Howard:1966}. In essence, computations can be
selected according to the expected improvement in decision quality
resulting from their execution. I. J.~\citet{Good:1968} independently
proposed using this idea to control search in chess, and later defined
``Type II rationality'' to refer to agents that optimally solve the
metalevel decision problem before acting. As interest in probabilistic
and decision-theoretic approaches in AI grew during the 1980s, several
authors explored these ideas further~\citep{Dean+Boddy:1988,Doyle:1988,Fehling+Breese:1988,Horvitz:1987b}.
Work by
\citet{Russell+Wefald:1988a,Russell+Wefald:1991a,Russell+Wefald:1991b}
formulated the metalevel sequential decision problem, employing an
explicit model of the results of computational actions, and applied
this to the control of game-playing search in Othello with encouraging
results.

An independent thread of research on metalevel control began with work
by \citet{Kocsis+Szepesvari:2006} on the UCT algorithm, which operates
in the context of {\em Monte Carlo tree search} (MCTS) algorithms.
In MCTS, each computation takes the form
of a simulating a randomized sequence of actions leading from a leaf of the
current tree to a terminal state. UCT is primarily a method for
selecting a leaf from which to conduct the next simulation.
It forms the core of the successful \textsc{MoGo} algorithm for Go playing \citep{Gelly+Silver:2011}.  It is situated 
within the theory of bandit problems \citep{Berry+Fristedt:1985}, applying the bandit algorithm
UCB1 \citep{Auer+et+al:2002} recursively to select actions to perform within simulations.

It is natural to consider whether the two independent threads are
consistent; for example, are bandit algorithms such as UCB1 approximate
solutions to some particular case of the metalevel decision problem
defined by Russell and Wefald? The answer, perhaps surprisingly, is no.
The essential difference is that, in bandit problems, every trial involves
executing a real object-level action with real costs, whereas in the metareasoning
problem the trials are {\em simulations} whose cost is usually independent of the
utility of the action being simulated. 
Hence, in UCT, bandit algorithms are applied to problems that are not bandit problems.

One consequence of the mismatch is that bandit policies are
inappropriately biased away from exploring actions whose current
utility estimates are low.  Another consequence is the absence of any
notion of ``stopping'' in bandit algorithms, which are designed for
infinite sequences of trials.  A metalevel policy needs to decide
when to stop deliberating and execute a real action.

Analyzing the metalevel problem within an appropriate theoretical
framework ought to lead to more effective algorithms than those
obtained within the bandit framework.  For Monte Carlo computations,
in which samples are gathered to estimate the utilities of actions,
the metalevel decision problem is an instance of the {\em selection
problem} studied in statistics~\cite{early-selection-refs}.  Despite
recent work by \citet{Frazier+Powell:2010}, the theory of selection
problems is less well understood than that of bandit problems.
Accordingly, we present in \secrefs{optimal-section}{context-section}
a number of results concerning optimal policies for the general case
and for specific sampling distributions. In particular, we give finite
bounds on the number of samples collected by optimal policies for
Bernoulli and Gaussian distributions. We also provide a simple
counterexample to the intuitive conjecture that an optimal policy
should not spend more on deciding than the decision is worth; in fact,
it is possible for an optimal policy to compute forever. We also show
by a simple counterexample that {\em index
policies}~\cite{Gittins:1989} do not necessarily exist for selection
problems.

Motivated by this theoretical analysis, we propose in \secrefs{approx-bayesian-section}{approx-nonbayesian-section}
two families of heuristic approximations, one for the Bayesian case and one for the distribution-free setting.
We show empirically that these rules give better performance than UCB1 on a wide range of standard (non-sequential) selection problems.
\secref{mcts-section} shows similar results for the case of guiding Monte Carlo tree search in the game of Go.