### 2012-03-30 paper/vago.uai/optimalpolicies.tex

%% \section{On optimal policies for selection}

%% Define selection problem and metalevel MDP representation.

%% \note{Unsure about terminology here: selection problems, sampling problems, 
%% metalevel probability model, metalevel decision problem (conflicts with Markov decision process?).}

%% ============================
%% \subsection{Selection problems}
%% ============================

%% \note{Formal definition of selection problems and the metalevel MDP with cost per sample (time value); also mention budgeted learning.}

In a \term{selection problem} the decision maker is faced with the choice between
alternative actions.  To make this choice, they may gather evidence about the
utility of each of these alternative, at some cost.  The objective is to maximize
the expected utility of the final action selected, less the cost of computation.

%% This problem has been studied in a number of related forms.  \note{Here}

%% The Bayesian ranking and selection problem \citep{Swisher+et+al:2003} is an important
%% special case where the computations performed are samples.

%%What the decision maker can know and what they can find out is formalized in:

\begin{dfn} \label{dfn:metalevel-model}
	A \term{metalevel probability model} is a tuple $M=(U_1,\dots,U_k,\Evidence)$ 
	consisting of jointly distributed random variables:
	\begin{itemize}
		\item Real random variables $U_1,\dots,U_k$, where $U_i$ is the utility of performing action~$i$ (called ``arms" in the bandit setting), and
		\item A countable set $\Evidence$ of random variables, each variable $E\in\Evidence$ being 
		      a computation that can be performed and whose value is the result of that computation.
	\end{itemize}
\end{dfn}
%% Not actually true that we necessarily define this with U's as roots:  in sampling models they are naturally /leaves/.

A metalevel probability model, when combined with a cost $c>0$ of computation,%
\footnote{The assumption of a fixed cost of computation is a simplification; 
	precise conditions for its validity are given by \citep{Harada:1997}.} 
defines a metalevel decision problem: what is the optimal strategy with which to choose a sequence 
of computations $E\in\Evidence$ to observe in order to maximize the expected utility 
of the action finally taken less the cost of computation?  Intuitively, 
this strategy should observe the variables which give the most evidence relevant
to deciding which action to perform, stopping when the cost of computation 
outweighs the benefit gained.

\begin{example}[Bernoulli sampling]\label{example:bernoulli}
%% One simple metalevel probability model is used as a model of the results of simulations \citep{someone}.
In the \term{Bernoulli metalevel probability model},
each action will either succeed or not $U_i\in\{0,1\}$, with an unknown latent frequency of success $\Theta_i$, 
and stochastic simulations of possible consequences $E_{ij}$ can be performed:
\begin{align*}
	\Theta_i &\iid {\rm Uniform}[0,1]                      &&\quad\text{for $i=1,\dots,k$}\\
	U_i \given \Theta_i &\sample {\rm Bernoulli}(\Theta_i) &&\quad\text{for $i=1,\dots,k$}\\
	E_{ij} \given \Theta_i &\iid {\rm Bernoulli}(\Theta_i) &&\quad\text{for $i=1,\dots,k$ and $j=1,\dots$}
\end{align*}

The \term{one-action Bernoulli metalevel probability model} has $k=2$,
$\Theta_1=m\in[0,1]$ a constant, and $\Theta_2\sample{\rm Uniform}[0,1]$.
\end{example}

%% Bounedness assumption
For simplicity, in the below we'll assume the utilities $U_i$ are bounded and so, 
without loss of generality, lie in $[0,1]$.

To formalize this problem, we use the theory of Markov Decision Processes (MDPs; 
see, for example, \citet{Puterman:1994}):

\begin{dfn}
A (countable state, undiscounted) \term{Markov Decision Process} (MDP) is a tuple $M=(S,s_0,A_s,T,R)$ where:
	$S$ is a countable set of states,
	$s_0\in S$ is the fixed initial state,
	$A_s$ is a countable set of actions available in state $s\in S$,
	$T(s,a,s')$ is the transition distribution 
	giving the probability of transitioning from a state $s\in S$ to a state $s'\in S$ after performing action $a\in A_s$,
	and $R(s,a,s')$ is the expected reward received when transitioning from $s\in S$ to $s'\in S$ after performing action $a\in A_s$.
\end{dfn}

\begin{dfn}\label{dfn:metalevel-mdp}
	Given a metalevel probability model%
		\footnote{Definition~\ref{dfn:metalevel-model} made no assumption about the computational result
			variables $E_i\in\Evidence$, but for simplicity in the following we'll assume that
			each $E_i$ takes one of a countable set of values.  Without loss of generality, 
			we'll further assume the domains of the computational variables $E\in\Evidence$ are disjoint.}
	$M=(U_1,\dots,U_k,\Evidence)$ and
	a cost of computation $c>0$, a corresponding \term{metalevel decision problem}
	is any MDP $M=(S,s_0,A_s,T,R)$ \begin{align*}
		S &= \{\bot\}\cup\{\langle e_1\dots, e_n \rangle : e_i\in E_i \text{ for all $i$,} \\
								& \qquad\qquad \text{for finite $n\ge0$ and distinct $E_i\in\Evidence$}\} \\
		s_0 &= \langle \rangle \\
		A_s &= \{\bot\}\cup\Evidence_s \\
	%
	\intertext{where $\bot\in S$ is the unique terminal state,
	where $\Evidence_s\subseteq\Evidence$ is a potentially state-dependent subset of allowed computations,
	and when given any $s=\langle e_1, \dots, e_n \rangle\in S$,
	computational action $E\in\Evidence$, 
	and $s'= \langle e_1, \dots, e_n, e \rangle\in S$ where $e\in E$, we have:}
	%
		T(s,E,s') &= P(E=e \given E_1=e_1,\dots,E_n=e_n) \\
		T(s,\bot,\bot) &= 1 \\
		R(s,E,s') &= -c \\
		R(s,\bot,\bot) &= \max_i \mu_i(s) % \max_i \IE[U_i \given E_1=e_1,\dots,E_n=e_n]
	\end{align*}		
	and where $\mu_i(s) = \IE[U_i \given E_1=e_1,\dots,E_n=e_n]$.
\end{dfn}

One can optionally add an external constraint on the number of computational actions, or their total cost,
in the form of a deadline (called a budget henceforth).  This bridges with the related area of budgeted learning \citep{Madani+et+al:2004}.
Although this feature is not formalized in the MDP, it can be added by including either time or past total cost 
as part of the state.

\begin{example}[Bernoulli sampling]\label{example:bernoulli2}
In the Bernoulli metalevel probability model (\exampleref{example:bernoulli}),
note that: 
\begin{align}
	\Theta_i \given E_{i1},\dots,E_{in_i} &\sim {\rm Beta}(s_i+1, f_i+1)  \label{eq:bernoulli1}\\
	E_{i(n_i+1)} \given E_{i1},\dots,E_{in_i} &\sim {\rm Bernoulli}\left(\frac{s_i+1}{n_i+2}\right) \label{eq:bernoulli2} \\
	\IE[U_i \given E_{i1},\dots,E_{in_i}] &= (s_i+1)/(n_i+2) \label{eq:bernoulli3}
\end{align}
by standard properties of these distributions, where $s_i=\sum_{j=1}^{n_i} E_{in_i}$
is the number of simulated successes of action $i$, and $f_i=n_i-s_i$ the failures.  By \eqref{eq:bernoulli1}, 
the state space is the set of all $k$ pairs $(s_i,f_i)$, \eqrefs{eq:bernoulli2}{eq:bernoulli3}
suffice to give the transition probabilities and terminal rewards, respectively.
The one-action Bernoulli case is similar, requiring only the two dimensional 
space $(s,f)$ giving the posterior over $\Theta_2$.
\end{example}

Given a metalevel decision problem $M=(S,s_0,A_s,T,R)$ one defines policies and 
value functions as in all MDPs.  A (deterministic, stationary) \term{metalevel policy} 
$\pi$ is a function mapping states $s\in S$ to actions to take in that state $\pi(s)\in A_s$.

The \term{value function} for a policy $\pi$ gives the expected total reward
received under that policy starting from a given state $s\in S$, and the \term{Q-function}
does the state when starting in a state $s\in S$ and taking a given action $a\in A_s$: 
\[
	V^\pi_M(s) = \IE^\pi_M\left[ \sum_{i=0}^N R(S_i,\pi(S_i),S_{i+1}) \given S_0 = s\right]	\label{eq:value-fn}\\
%%	Q^\pi_M(s,a) &= \IE^\pi_M\left[ \sum_{i=0}^N R(S_i,\pi(S_i),S_{i+1}) \given S_0 = s, A_0 = a\right] \label{eq:qvalue-fn}
\]
where $N\in[0,\infty]$ is the random time the MDP is terminated, i.e.,
the unique time where $\pi(S_N)=\bot$%
\footnote{Whenever $N<\infty$, $S_{N+1}=\bot$ the unique terminal state.
Instead of having a random termination time, one can equivalently make
the state $\bot$ absorbing with zero reward on all transitions.}, 
and similarly for the Q-function $Q^\pi_M(s,a)$.

As usual, an \term{optimal policy} $\pi^*$, when it exists, is one which maximises 
the value from every state $s\in S$, i.e., if we define for each $s\in S$
\[
	V^*_M(s)   = \sup_\pi V^\pi_M(s),
%%	Q^*_M(s,a) &= \sup_\pi Q^*_M(s,a),
\]
then the optimal policy $\pi^*$ satisfies $V^{\pi^*}_M(s) = V^*_M(s)$
for all $s\in S$, where we break ties towards stopping $\bot$.
%% and by standard results, $\pi^*(s) \in \argmax_{a\in A_s} Q^*_M(s,a)$, where

The optimal policy must balance the cost of computations with the improved decision
quality that results.  This is made clear by giving by expressing the value function:

\begin{thm}\label{thm:value-of-computation}
	The value function of a metalevel decision process $M=(S,s_0,A_s,T,R)$ is of the form
	\[
		V^\pi_M(s) = \IE^\pi_M[ -c\,N + \max_i \mu_i(S_N) \given S_0=s]
	\]
	where the random variable $N$ denotes the total number of computations performed,
	and similarly for $Q^\pi_M(s,a)$.
\end{thm}

\begin{proof}
	Follows immediately from \eqref{eq:value-fn} and the definition of the
	reward function in \dfnref{dfn:metalevel-mdp}.
\end{proof}

In many problems, including the Bernoulli sampling model of \exampleref{example:bernoulli2},
the state space is infinite.  Does this preclude solving for the optimal policy?  Can 
infinitely many computations be performed?

There is in full generality an upper bound on the \emph{expected} number of computations
a policy performs:

\begin{thm}\label{thm:bounded-expected-computations}
	The optimal policy $\pi^*$'s expected number of computations is bounded by the 
	perfect value of information \citep{Howard:1966} times the inverse cost $1/c$:
	\[
		\IE^{\pi^*}[N\given S_0=s] \le \frac{1}{c} \left(\IE[\max_i U_i\given S_0=s] - \max_i \mu_i(s)\right).
	\]
	Further, any policy $\pi$ with infinite expected number of computations 
	$\IE^\pi[N]=\infty$ has negative infinite value, so in particular the optimal
	policy stops with probability 1.
\end{thm}

\begin{proof}
	The first follows as in state $s$ the optimal policy has value at least that
	of stopping immediately $\max_i \mu_i(s)$, and as $\IE \max_i\mu_i(S_N) \le \IE \max_i U_i$ by Jensen's inequality.
	The second follows from \thmref{thm:value-of-computation} as $c>0$.
\end{proof}

However, although the \emph{expected} number of computations is always bounded,
there are important cases in which the \emph{actual} number is not, such as
the following example inspired by the sequential probability ratio test \citep{Wald+1945}:

\begin{example}\label{example:sprt}
Consider the Bernoulli sampling model for two actions but with different priors:
	$\Theta_1=1/2$,
	and $\Theta_2$ is $1/3$ or $2/3$ with equal probability.
%
Simulating action 1 gains nothing, and after $(s,f)$ simulated successes and failures of action 2
the posterior odds is
\[
	\frac{P(\Theta_2=2/3\given s,f)}{P(\Theta_2=1/3\given s,f)} = \frac{(2/3)^s(1/3)^f}{(1/3)^s(2/3)^f}= 2^{s-f}.
\]
Note that these odds completely specify the posterior distribution of $\Theta_2$,
and thus the distribution of the utility and all future computations.  Thus, whether
it is optimal to continue is a function only of this posterior odds, and thus $s-f$.
%% In particular,
%% note that after equal numbers of successes and failures, $s-f=0$ and so the posterior 
%% distribution over $\Theta_2$ is equal to the prior.
For sufficiently low cost, the 
optimal policy samples when $s-f$ equals $-1$, $0$, and $1$.  But with probability
$1/3$, a state with $s-f=0$ transitions to another state $s-f=0$ after two samples, 
giving finite, although exponentially decreasing, probability to arbitrarily long 
sequences of computations.
\end{example}

Fortunately, in a number of settings, including the original Bernoulli model of \exampleref{example:bernoulli},
we can prove an upper bound the number of computations.  For reasons of space,
and for its later use in \secref{sec:blinkered}, we prove here the bound for the one-action Bernoulli model.

Before we can do this, we need to get an analytical handle on the optimal policy.
The key is through a natural approximate policy:

\begin{dfn}\label{dfn:myopic}
	Given a metalevel decision problem $M=(S,s_0,A_s,T,R)$,
	the \term{myopic policy} $\pi^m(s)$ is defined to equal $\argmax_{a\in A_s} Q^m(s,a)$ 
	where $Q^m(s,\bot) = \max_i \mu_i(s)$ and
	\begin{equation}\label{eq:myopic}
		 Q^m(s,E) = \IE_M[ -c + \max_i \mu_i(S_1) \given S_0 = s, A_0 = E].		
	\end{equation}
\end{dfn}

The myopic policy takes the best action, to either stop or perform a computation,
under the assumption that at most one further computation can be performed.
%% This policy is typically easy to compute, although not necessarily very good.

It does have importance, however, in how it constrains the optimal policy in \thmref{thm:optimal-myopic}
and its partial converse \thmref{thm:myopic-optimal}:

%% \note{Theorem: if Optimal stops in x, myopic stops in x (converse is more useful!)}

\begin{thm}\label{thm:optimal-myopic}
	Given a metalevel decision problem $M=(S,s_0,A_s,T,R)$
	if the myopic policy performs some computation in state $s\in S$,
	then the optimal policy does too, i.e., if $\pi^m(s)\neq\bot$ then $\pi^*(s)\neq\bot$.
\end{thm}

\begin{comment}
	\begin{proof}
		Observe that the myopic Q-function \eqref{dfn:myopic} is equivalently given by
		\[
			Q^m(s,a) = Q^\bot(s,a)
		\]
		where $\bot$ is the policy which immediately stops $\bot(s)=\bot$.
		Thus $Q^m(s,a) \le Q^*(s,a)$.  If the optimal policy stops in a state $s\in S$ then
		\[
			Q^{\pi^*}(s,a) \le \max_i \mu_i(s),
		\]
		and so the same holds for $Q^m$, showing the myopic stops.
	\end{proof}
\end{comment}

% Similarity to an optimal stopping result
	% Search <blah> for one-step lookahead rules.
	% But HERE is generalized to a stopping problem where there are many ways to continue.

This shows that the myopic policy is conservative in stopping, so stopping too early
is one of its approximation errors.  However myopic stopping can be used to control optimal stopping:

\begin{dfn}\label{dfn:closed}
	Given a metalevel decision problem $M=(S,s_0,A_s,T,R)$,
	a subset $S'\subseteq S$ of states is \term{closed under transitions}
	if whenever $s'\in S'$, $a\in A_{s'}$, $s''\in S$, and $T(s',a,s'')>0$,
	we have $s''\in S'$.
\end{dfn}

\begin{thm}\label{thm:myopic-optimal}
	Given a metalevel decision problem $M=(S,s_0,A_s,T,R)$
	and a subset $S'\subseteq S$ of states closed under transitions,	
	if the myopic policy stops in all states $s\in S$
	then the optimal policy does too.	
\end{thm}

\begin{comment}
	\begin{proof}
	Take any $s^*\in S'$, and note that all states the chain can transition
	to from $s^*$ are also in $S'$, by transition closure.  Defining $m(s) = \max_i\mu_i(s)$, 
	observe the myopic stopping for all such states implies that
	\begin{align*}
		\IE^{\pi}[(m(S_{j+1}) - c)\, 1(j<N)\given S_0=s^*] \\
		\le \IE^{\pi}[m(S_{j})\, 1(j<N)\given S_0=s^*]
	\end{align*}
	holds for all $j$, and as a result:
	\begin{align*}
		V^\pi(s) 
		&= \IE^{\pi}[ - c N + m(S_N) \given S_0=s^*] \\
		&= \IE^{\pi}[m(S_0) + \sum_{j=0}^{N-1} (m(S_{j+1}) - c - m(S_j)) \given S_0=s^*] \\
	%%	&\le \IE^{\pi}[m(S_0) + \sum_{j=0}^{N-1} 0 |S_0=s] \\		
		&\le \max_i\mu_i(s^*) \qedhere
	\end{align*}
	\end{proof}	
\end{comment}

Using this, we can prove our bound:

\begin{thm}\label{thm:one-action-bound}
	The one-action Bernoulli decision process with constant arm $m\in[0,1]$ 
	performs at most $m(1-m)/c-3 \le 1/4c-3$ computations.
\end{thm}

\begin{comment}
	\begin{proof}
		By \dfnref{dfn:myopic} and \exampleref{example:bernoulli2}, the myopic policy stops in
		a state $(s,f)$ when
		\begin{align}
			c \ge &\mu\max(\mu^+,m) + (1-\mu_i)\max(\mu^-, m) - \max(\mu,m) \label{eq:stopping}
		\end{align}
		where $\mu=(s+1)/(n+2)$  is the mean utility for action~2, where $n=s+f$,
		$\mu^- = \mu - \mu/(n+3)$, and $\mu^+ = \mu + (1-\mu)/(n+3)$
		are the posterior means of action 2 after simulating a failure and a success,
		%% and $\mu^-_i=(s_i+1)/(s_i+f_i+3) $ and $\mu^+_i = (s_i+2)/(s_i+f_i+3)$
		%% are the posterior means of action $i$ after simulating a failure and a success,
		respectively.  Whenever \eqref{eq:stopping} holds, stopping is preferred to sampling.

		Fixing $n$ and maximizing over $\mu$, we get sufficient condition for stopping
		\begin{equation}
			c \ge \frac{m(1-m)}{(n_i+3)} \qquad\qquad   n_i\ge \frac{m(1-m)}{c} - 3  \label{eq:myopic-stopping}
		\end{equation}
		Since the set of states satisfying \eqref{eq:myopic-stopping} is closed under
		transitions ($n$ only increases), by \thmref{thm:optimal-myopic}.  Finally, note $\max_{m\in[0,1]} m(1-m)=1/4$.
	\end{proof}	
\end{comment}

A key implication is that the \emph{optimal} policy can be computed
in time quadratic in the inverse cost $O(1/c^2)$.  This is particularly appropriate when the cost of 
computation is relatively high, such as in simulation experiments \citep{Swisher+et+al:2003},
or when the decision to be made is critical.