### 2012-06-10 paper/vago.ecai/case-for-support.tex
fixes

\documentclass[11pt]{letter}
\usepackage{eurosym}

\signature{David Tolpin}

\begin{document}

\begin{letter}{}
\opening{Dear ECCAI Travel Grant Officer,}

I, David Tolpin, am a full-time PhD student at the Department of
Computer Science at Ben-Gurion University of the Negev, Israel. For the last
three years, I have been researching {\it metareasoning in
  problem-solving search}, a field of Artificial Intelligence seeking
systematic approaches to development of heuristic algorithms for hard
problems. My and my advisor's joint paper, {\bf VOI-aware MCTS},
presenting a result of ongoing research, has been accepted to
ECAI~2012 for poster presentation.

The paper addresses an important issue of simple regret minimization
in Monte Carlo sampling in trees, and presenting the paper at the
conference and getting feedback from other researchers,
as well as attending other presentations in this and other fields of
AI is important for my studies in AI. However, travel,
lodging, and registration costs are high, and without financial 
support I will have difficulty attending the conference.

The anticipated expenses are :

\begin{itemize}
\item registration --- \euro{120}
\item accomodation --- \euro{180} (student room)
\item airplane tickets --- \euro{680} (current lowest available rate
  from the local travel agent)
\end{itemize}

{\bf Total: \euro{980}}

As the expenses exceed the ECCAI travel grant, the Computer Science
department agreed to provide additional funds (see the letter from the 
department).

\closing{Sincerely Yours,}

\end{letter}
\end{document}